---
title: "Project 1"
author: "Jose Singer-Freeman"
date: "2023-06-25"
output: github_document
---

## I. Introduction

This  project for Stat 558 consists of a vignette where I show how to retrieve certain information from the polygon.io Market Data APIs.  After retrieving the data from the polygon.io API, I will process it and preform a bit of  exploratory data analysis.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)

library(httr)
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(knitr)
library(scales)
rm(list=ls())

```


## II. Required Packages

The packages needed to run the code to create this vignette are the following:  

- jsonlite
- httr  
- tidyverse
- ggplot2
- knitr
- scales


## III. Background  

### What will we look at?

1.  First, we will retrieve price and transaction information for all publicly traded equities for a particular date of your choice.  The endpoint for these data is "Aggregate (Bars)" under the Stocks API.


2.  Among these equities, we will choose the tickers with the 5 highest trading transactions on that date. Equities  may include stocks, warrants, ETFs and publicly traded trusts.  

3.  After finding the top 5 equities, we will retrieve price/transaction information for them for a period of your choice. We will also obtain information about the issuer behind these tickers.  The endpoints for the price/transaction and issuer information are Grouped Daily (Bars) and Tickers under the Stocks API.  


###  What inputs will you need/be able to enter?

  - The date for determining the 5-highest transacted equities
  - The range of dates for the historic information on those 5 equities
  
### Why multiple endpoints?

The  polygon.io API is designed to query all available financial information on all equity  securities for a single day or all available financial information on a single equity security for a range of dates.  It does not have a single query to get both. Also, the issuer information for a ticker is maintained in yet another endpoint. 


## IV. Functions to Retrieve Data

`find_top_5`

This function returns the 5 equities with highest number of transactions for a particular data, including the following information:
  - ticker
  - volume-
  - weighted price
  - open price
  - close price
  - high price,
  - low price
  - number of transactions 
  - date  (yyyy-mm-dd)
  - intra-day price change  (calculated variable) 
  -  gain (1=yes/0=no)(binary)(calculated variable)
  


```{r}
#get info from API

find_top_5<-function(date){
base<-'https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/'
adjusted<-'?adjusted=true'
info_key<-'&apiKey=7FTp6uzAjKcWUI1BJEO7WppG3wyOZOy9'
API_URL<-paste0(base,date, adjusted, info_key)
daily_stock_info<-GET(API_URL)
df_daily_stock<-fromJSON(rawToChar(daily_stock_info$content))

#Re-format date/time of selected date (the format provided by the API is number of seconds since 1/1/1970) and relabel variables. 
df_daily_stock_results<-df_daily_stock$results%>%
  dplyr::mutate(date=as.POSIXct(t/1000, origin="1970-01-01"))%>%
  dplyr::rename(ticker=T, volume=v, weighted_price=vw, open_price=o, close_price=c, high_price=h, low_price=l, num_transactions=n)%>%
  select(-t)

#Pick the top 5 tickers with highest number of transactions in the chosen date 
#Also create a variable for the change in price for the day.
#Remove hours/minutes from date

top_5<-df_daily_stock_results%>%
  mutate(price_change=round(close_price/open_price-1,3), Gain=if_else(price_change>0,1,0), date=strftime(date, format="%Y-%m-%d"))%>%
  arrange(desc(num_transactions))%>%
  slice(seq(5))

return(top_5)
}
```

`get_company_info`  

This function returns price/transaction information and issuer information for the 5 top issuers for 1 day.  To do so, it uses another function `get_1_company`, which returns this information for 1 issuer .


```{r}

# Function to get name of issuer, number of employees, outstanding shares for the top 5 tickers:

get_1_company<-function(xyz){
  #generate URL for the API for 1 ticker
  ticker<-xyz
  base<-'https://api.polygon.io/v3/reference/tickers/'
  info_key<-'?apiKey=7FTp6uzAjKcWUI1BJEO7WppG3wyOZOy9'
  API_URL<-paste0(base,ticker,info_key)
  #connect to API
  ticker_info<-GET(API_URL)
  #Convert retrieved information (JSON format) to an R list 
  df_ticker<-fromJSON(rawToChar(ticker_info$content))

  #Form a vector from the various variables for the 1 ticker 
df_tick_result<-cbind(
  ticker=df_ticker$results$ticker_root,
  company=df_ticker$results$name,
  outstanding=df_ticker$results$share_class_shares_outstanding,
  employees=as.numeric(df_ticker$results$total_employees))
return(df_tick_result)
}

get_company_info<-function(top_5){

#create vector with top 5 ticker names
top_5_tickers<-top_5$ticker
  
#Initialize dataframe that will store information for 5 companies 
df_ticker_res<-data.frame()

#loop through the 5 tickers to invoke the function that obtains the issuer information for each ticker

for (i in 1:5){
  oo<-data.frame(get_1_company(top_5_tickers[i]))
  df_ticker_res<-dplyr::bind_rows(df_ticker_res, oo)
  }


##MERGE the financial information with issuer information
top_5_all<-left_join(top_5,df_ticker_res, by="ticker")%>%mutate(market_cap=as.numeric(outstanding)*weighted_price/1000000000)
#replace employees=NULL with 0s
top_5_all$employees[is.na(top_5_all$employees)] <- 0

return(top_5_all)
}

```



```{r}

#HISTORICAL INFO FOR EACH EQUITY
#Function to obtain prices and transactions for one ticker for a period of several days from the Polygon API


# Get info from API

get_daily_info<-function(tick, daterange){
base<-'https://api.polygon.io/v2/aggs/ticker/'
ticker<-tick
range<-'/range/1/day/'
dates<-daterange
adjusted<-'?adjusted=true&sort=asc&limit=120'
info_key<-'&apiKey=7FTp6uzAjKcWUI1BJEO7WppG3wyOZOy9'
API_URL<-paste0(base,ticker, range, dates, adjusted, info_key)
aggr_ticker_info<-GET(API_URL)
df_ticker_aggr<-fromJSON(rawToChar(aggr_ticker_info$content))

df_aggr_ticker_results<-as_tibble(df_ticker_aggr$results)%>%
  mutate(ticker=df_ticker_aggr$ticker)
return(list(df_aggr_ticker_results))
}

#Function to  obtain the daily info for each of the 5 top equities by invoking get_daily_info function above. 

get_history<-function(daterange, top_5_tickers){
  
#initialize data frame
df_daily<-data.frame(v=integer(),
                     vw=double(),
                     o=double(),
                     c=double(),
                     h=double(),
                     l=double(),
                     n=double(),
                     t=double())

for (i in 1:5){
  oo<-data.frame(get_daily_info(top_5_tickers[i], daterange))
  df_daily<-dplyr::bind_rows(df_daily, oo)
    }

df_daily<-df_daily%>%
  dplyr::rename(volume=v, weighted_price=vw, open_price=o, close_price=c, high_price=h, low_price=l, num_transactions=n)%>%
  dplyr::mutate(date=as.POSIXct(t/1000, origin="1970-01-01"),price_change=scales::percent(round(close_price/open_price-1,3)))%>% 
  dplyr::select(-t)%>%
  dplyr::relocate(ticker)

return(df_daily)

}

```


`wrapper`
This function calls the other functions and yields two data sets:  one for the initial date (financial and issuer information) and another for the range of dates ( financial information only)

```{r}
wrapper<-function(x, y){
top_5<-find_top_5(x)
top_5_all<-get_company_info(top_5)
top_5_tickers<-top_5$ticker
df_daily<-get_history(y, top_5_tickers)
return(list(top_5_all, df_daily))
}

#CALL THE WRAPPER FUNCTION.  
#USER -  PLEASE PROVIDE THE DATE AND DATE RANGE

date<-'2023-05-19'
daterange<-'2023-01-01/2023-05-01'

h<-wrapper(date, daterange )
top_5_all<-h[[1]]
df_daily<-h[[2]]

```



```{r echo=FALSE}


t<-unlist(strsplit(daterange, "/"))
first_date<-format(as.Date(t[1]),format="%B %d, %Y" )
second_date<-format(as.Date(t[2]),format="%B %d, %Y" )
```


## Exploratory Data Analysis

The top 5 most traded equity securities on `r format(as.Date(t[1]),format="%B %d, %Y")` are:

```{r echo=FALSE}

df_ticker_res<-top_5_all%>%dplyr::select(ticker, company, outstanding, employees)%>%mutate(outstanding=round(as.numeric(outstanding)/1000000000,2), employees=prettyNum(employees, big.mark=",", scientific=FALSE))%>%arrange(ticker)

knitr::kable(df_ticker_res, col.names = c("ticker","issuer/equity","outstanding shares ($bn)","number of employees"))

f<-paste0("top traded equities on ",format(as.Date(date),format="%B %d, %Y"))  # This is the title for the graph that follows

ggplot(data=top_5_all, aes(x=market_cap, y=as.numeric(employees), label=ticker))+geom_point()+geom_text(hjust = -0.1, nudge_x = 0.05)+labs(title=paste0("Market Cap v. Number of Employees for ",f), y="employees", x="market capitalization ($bn)")+scale_x_continuous(label=comma)+scale_y_continuous(label=comma)  

```

```{r echo=FALSE}
#Plot the daily price increase for each of the 5 stocks with highest number of transactions 

ggplot(data=top_5_all, aes(x=ticker, y=price_change, fill=Gain))+geom_bar(stat="identity",show.legend = FALSE)+labs(title=paste0("percentage gain/loss for ",f), y="Price Increase")


```

```{r echo=FALSE}

ggplot(df_daily, aes(y=weighted_price))+geom_boxplot()+labs(title=paste0("Weighted Prices for ", format(as.Date(date),format="%B %d, %Y",y="weighted price")))+facet_wrap(ticker~.)


```

```{r echo=FALSE}
ggplot(df_daily, aes(x=date, y=close_price))+geom_line()+facet_wrap(ticker~.)+labs(y="close price", title=paste0("Close prices for ",first_date," through ", second_date))
```



